Revisiting 'Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation

### 2023-06-16 analysis/article/paper.qmd
rewrite transparency chapter

---
title: Revisiting 'Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  I revisit the dynamic public goods game developed by Gächter et al. (2017) to study cooperation under dynamic interdependencies. Collecting data from both a convenient (students) and an inconvenient (general population) sample, I not only reproduce some of the authors' original observations but also test their novel game's generalizability. Appending a charitable dictator game, I find no correlations between behavior in the charitable context and the dynamic game. This applies to students and the general population sample alike. Because the study of inexperienced general population samples raises methodological challenges, such as fatigue and dropouts, this research approaches them. Doing so, I provide simple solutions to run reliable interactive experiments online.

@@

Today's actions are tomorrow's result. There are many settings in which current decisions affect future outcomes and with it, future decision spaces. 
<!-- Think of lifetime consumption‐smoothing and save more tomorrow -->
Opting for environmental friendly policies today not only reduces carbon dioxide omissions immediately but also helps us to reach the Paris climate targets tomorrow. Deferring these policies, may not necessarily prevent us from reaching these targets, but it requires more effort in the future compared to a path that includes immediate action @HaenselEtAl2022. Aiming at certain goals, today's actions (or the omission thereof) not only affect intermediate outcomes but also the number of paths one can choose from that lead to that specific goal.

@@

A game designed by Gächter, Mengel, Tsakas & Vostroknutov [-@GMTV2017 hereafter, GMTV] as well as Stefan Große (unpublished) shows that it is fairly simple to add this element of realism. They incorporate interdependencies into a _dynamic_ public goods game (dPGG) by defining endowments as the income of previous periods. Consequently, participants' actions in a given period affect their number of actions in subsequent periods: the more (less) they earn now, the more (less) they can contribute in the next period. Importantly, this modification qualitatively yields the same rational predictions as the static game (i.e. free-riding and the under-provision of the public good). It is thus, equally well suited to study dilemma situations.

Because there is surprisingly little experimental research on interdependencies^[One exemption are @Moser2019, who built on GMTVs' design to investigate leadership.], I reproduced one of GMTVs' treatments to compare dynamics across (in)experienced samples to investigate its generalizability.
@GKLS2020 find that static public goods games do not generalize well to real-world climate action. They also find that generalizability depends on structural resemblance of the public goods game with the context of climate change mitigation: Greater resemblance improves generalizability. Because GMTVs' dynamic setting has a more realistic property—namely, interdependencies—one would expect it to be better suited to inform public policy. To test this intuition, I not only ran the experiment with different samples but also
<!--, that is, with participants who make their decisions in an environment which is more natural than the lab.-->
observed the participants' behavior in voluntary climate actions (VCA).^[The VCA is a (charitable) dictator game where each participant is a dictator dividing her budget between herself and some organization linked to the reduction of CO2 emissions. @EckelGrossman1996 were the first to implement a charitable dictator game observing contributions of 30% of the endowments. Like @GKLS2020, @CarpenterEtAl2008 report that students make lower contributions to charity than community members.] This yielded a setting similar to the one of @GKLS2020, which allows me to analyze how behavior in the abstract game translates into real-world action across samples. This research shows that the dynamic setting _does not_ add any advances with respect to generalizability of results.

@AGM2018 conducted static public goods games in the lab and on MTurk to draw (and report) lessons from online experimentation. This study extends this literature [see, e.g., @SnowbergYariv_2021; @GuptaRigottiWilson_2021; @GoodmanPaolacci2017; @AmirEtAl2012] by focusing on an _inconvenient_ sample (i.e. one that is completely inexperienced sample that has not been exposed to interactive experiments before) playing a computationally more complex game. This required me to design a robust (and thus, more complex) software to minimize attrition. I collected paradata---which captures screen time and survey navigation, for instance---to assess the desired fluency and feasibility of the experiment. Like @AGM2018, this research is of practical relevance as it reports on the robust design that made the dynamic game feasible.

@@

After commenting on transparent research practices and reporting the methods in @sec-transparency and @sec-methods, this paper is organized along these findings: The confirmatory reproduction is reported in @sec-replication, the online feasibility in @sec-feasibility and the generalizability in @sec-generalizability. @sec-conclusion concludes. Because I observed some (qualitatively inconsequential) differences in reproducing GMTVs' analysis using their data, I report these in the [Appendix](@sec-appendix).

@@

Well before credibility crises comprised a variety disciplines, @bemwriting [p. 2] indirectly conceded discrepancies between fuzzy research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations were established to tie the researchers' hands when doing confirmatory research--reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b). Today, researchers can not only choose where to pre-register but also decide about the level of detail they provide. More severely, nothing stops researchers from pre-registering multiple hypotheses opposing each other [see, e.g., @SimmonsEtAl2021; and @PhamEtAl2021 for a discussion].

@WaldronAllen2022 report a wide variety of depth of pre-registrations in the field of Neuropsychopharmacology and document that the _'pre-registered'_ label can be obtained with minimal effort. In absence of transparency and scrutiny pre-registrations run into danger to miss the target of making science more credible. Even more so, they may have adverse effects as the _'pre-registered'_ label becomes cheap (too) to acquire.

Discussing the costs and benefits of pre-registrations, @Olken2015 stresses the uncertainty of research processes and promotes moderate approaches. As @WaldronAllen2022, he argues that pre-registrations not necessarily eliminate  flexibility. These 'moderate' or 'flexible' approaches require transparency to reduce concerns of _'data-mining'_ [@Olken2015 p. 61] by showing (a) the analyses I planned to run, (b) the analysis that made most sense after collecting the data and the transition from (a) to (b). Because this is neither incentivized nor easy to verify, we need tools that make transparency convenient to establish and to scrutinize.
I propose version control (such as Git) as one such tool and illustrate its application on this paper:  GitHub is a website and cloud-based service where developers---and researchers alike---can store and manage their code. The service is based on Git^[Git is a specific open-source version control system developed in 2005.] and designed for version control. Importantly, changes are time-stamped and can be tracked. Moreover, one can create branches, that is, duplicates of code either to work collaboratively or to archive a certain state.

To sum up, pre-registrations alone are necessary but not sufficient to restore the credibility in experimental sciences. Pre-registrations must come with additional emphasis on transparency covering the whole research process as well as with a higher level of scrutiny that looks beyond mere labels.

In fact, the experiment reported in this article qualifies for the _'pre-registered'_ label at a first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run when I designed the experiment on [GitHub](https://shorturl.at/bAGKY).^[https://shorturl.at/bAGKY to run the code, you need to execute the .Rmd files in this repository in the order that is indicated by file names.]

At a second glance, the analysis that made the most sense after having collected the data and that is reported here is a little different though: Because the more representative subject pool was exhausted earlier than expected, I recruited students which opened up a new research direction: assessing generalizability. Hence, this article combines both confirmatory as well as exploratory research.

Accordingly, I edited many of the analyses scripts.^[The source code of this document contains the analysis and can be found [here](https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd): https://shorturl.at/ilpAG] However, because the originally planned analysis code is archived and reproducible, these changes are transparent. Importantly, I followed a literate programming approach [@Knuth_1984; @AkhtarYe_2023]. Hence, all documents needed to analyze the data and to write the report stem from the same source. This establishes consistence between the commands one tells the statistical software to do and the explanation one tells human beings one told the statistical software to do. As such, (a), (b) and the transition from (a) to (b) are not only transparent but also comprehensible.

<!--
I archived detailed analyses scripts and created a [branch](https://github.com/Howquez/coopUncertainty/tree/July21Replication) a few days _before_ we collected data and pre-registered the analysis via the American Economic Association's RCT Registry [@preregistration].
-->

@@

In the terminology of @Hamermesh2007, I ran both a _pure_ as well as a _scientific_ reproduction^[@Parsons2022 use the term of a _conceptual replication_ which means the same.] of one treatment of GMTVs' dynamic public goods game. The pure reproduction re-analyzes the original data. [Appendix A](#A:-Pure-Replication) documents [errors I identified]() in the original paper. The scientific reproduction, where I utilize a different sample drawn from a different population in a different situation, is described in the following sections.

@@

Considering the design, the only differences between such a _static_ game (with partner matching) and GMTVs' design are _dynamics_: Instead of receiving fresh endowments every period, participants receive one endowment only at the beginning of the first period. A participant's endowment in the second period is the wealth she accumulated in the first period. A participant's endowment in the third period is the wealth she accumulated in the first two periods. And so on. Hence, a decision in one period has consequences on future endowments and, ultimately, growth paths. For this reason, the game is described as a _dynamic_ public goods game.

@@

Like GMTV I employ a real giving task after the abstract game. Specifically, I employed a (charitable) dictator game where each participant is a dictator dividing her budget between herself and some organization. Whereas GMTV chose _Doctors without Borders_ as a receiver, I chose another organization in another context: like @GKLS2020, I employed a VCA, where participants could donate any amount of their earnings to offset carbon dioxide (that is, retire emission permits from the EU ETS).^[Importantly, @GKLS2020 made the VCA decision with a fresh endowment _before_ they played the abstract game. I deviate from their procedure to match GMTVs' procedure.]
To ensure that each participant had the same basic level of information about the impact of their decision, I provided some basic information about the mechanism. The information also highlighted that the mitigation came into operation on a European level.
Finally, I informed the participants that the documentation of individual and aggregate contributions were to be posted immediately after the concluding the sessions online. To avoid privacy or social image concerns, participants learned their unique and random IDs, which they needed to identify their individual contributions.
The document certified that their contributions have been used to offset [1.82 tons](https://www.compensators.org/compensatelist/?searchterm=stefan+traub) of carbon dioxide emissions.

@@

I recruited the participants from the so called _[HamburgPanel](https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html)_ using HROOT [@hroot]. The panel is provided by the University of Hamburg's Research Laboratory, which used a randomized last digits approach to build the panel while drawing from the population of citizens of Hamburg, Germany. Because the sample was exhausted at one point, I also recruited students from the University of Hamburg.

@@

Throughout this paper, I will compare results of my experiment with the results of GMTVs' NOPUNISH 10 Period treatments. I am thus, referring to three different samples utilized at two points in time: the University of Nottingham's students (in late 2012), Hamburg's citizens and the University Hamburg's students (both in July 2021). @tbl-sample-properties gives an overview.

@@

The experiment was logistically complex for several reasons. First, the sample was inexperienced. Second, the experiment was interactive and synchronous. Third, the underlying game was dynamic and interdependent. This makes dropouts not only more likely but also more expensive, which is why attrition was a major concern implementing the experiment.

I chose oTree [@oTree] to implement the experiment because it is open-source, well documented and very flexible. Its [Bootstrap](https://getbootstrap.com/) (a powerful frontend toolkit) integration allowed me to make the graphical user interface interactive, appealing and easy to navigate. The [Highcharts library](https://www.highcharts.com/) made it easy to visualize results and to communicate dynamics. Insofar, oTree served a good tool to enhance the participants' user experience and thus, to make dropouts less likely.

Which features were required to handle dropouts? First, participants had to be matched to form a group _after_ comprehension questions were answered successfully. Importantly, participants were grouped by the order they answered these questions to reduce waiting times. While waiting for other players to form the group, the participants saw a wait-page informing them that they are waiting for other participants to arrive and that they do not have to wait for longer than 10 minutes. The screen also informed them that they would receive a _patience bonus_ of one Euro after the expiration of that time (or what was left of it). Second, participants only had 10 minutes to make the first contribution and 4 minutes for the remaining contributions. After this time expired, participants were replaced by bots that made random contributions. In this case, the remaining group members were informed about the replacement. Both features were implemented to limit wait times and boredom for other participants. @sec-feasibility shows that the first feature became effective in some cases, wheres the second feature did not.

@@

Participants entered the experiment at appointed times remotely from home. They first saw a welcome screen. After agreeing to the privacy policy, they could proceed to the instructions individually. Having read these instructions, each participant has also seen a demo-screen explaining the user interface. Before proceeding, they had to answer six comprehension questions correctly to avoid confusion in later stages [@FerraroVossler2010]. Subsequently, they saw a waiting screen until they could be matched with three other participants, who have answered the comprehension questions correctly. Once matched, they were exposed to the decision screen over ten periods. At the end of the last period, participants saw results of all periods. Subsequently, they made their VCA decision, before I elicited risk preferences [@HoltLaury2002] and finished with GMTVs' questionnaire.

@@

The experiment lasted around `r total_completion_time` minutes on average. The earnings  averaged `r earningsMean %>% round(digits = 2)` Euros (sd = `r earningsSD  %>% round(digits = 2)`).^[This values include earnings from the incentivized risk elicitation task that is not part of the analysis.]

@@

## Pre-registered GMTV Reproduction {#sec-replication}

@@

First, I ask whether the samples differ with respect to their initial contributions to the public good. Is the reproduction sample (consisting of both students and non-students) more pro-social than the original sample? A two-sided rank sum test reveals that it is not (p=`r rs1`). Both samples contributed 10 tokens, that is, 50% of their endowments on average (median and mean). Moreover, both samples' initial contributions resemble initial contributions participants usually make in the static game with partner matching.^[See Figure 3B in @fehrgaechter2000 [p.989], for instance.] However, in the dynamic game presented here, I are particularly interested in the subsequent periods because differences add up exponentially. Do the two groups remain similar over the course of time?

@@

In particular, do the two samples' contributions follow the same path over the 10 periods they played? The answer is _no_. @fig-share-of-contributions illustrates that the samples make similar contributions at the beginning and the end of the game but behave differently in between. More precisely, the left panel--depicting the average contributions in absolute terms--shows that the original sample contributed more than the reproduction sample _in all but the first and last period_.^[In period five, this difference is significant: a two-sided rank sum test yields p=`r rs2`. ] For this reason, the original sample's behavior differs from the reproduction sample's behavior in two aspects: it contributes more and exhibits a considerable drop in the last period (whereas the reproduction sample's contributions flatten).

@@

Again, both samples' behavior resembles the contributions participants usually make in the static game with partner matching: contributions equal approximately half of endowments in the very first period and decrease to around ten percent of endowments by the last period.^[The right panel is thus, comparable to the visualizations _and results_ in the static game. See, for instance, Figure 1B in @fehrgaechter2000 [p.986].] In the dynamic game presented here, however, different paths lead to different levels of wealth -- even if they share the same start- and end-points. I am thus, more interested in the contributions' implications for wealth generation and growth.

@@

How do the different contribution-paths translate into wealth?^[To measure wealth and growth, I define a variable called _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed).] Given that the original sample contributed more, one would expect the respective groups to be  more wealthy. A mere mean comparison indicates just that: An average group in the original sample accumulated about `r round(originalWealth)` tokens. In contrast, an average group in the reproduction sample accumulated about `r round(replicationWealth)` tokens. This difference is insignificant at conventional levels though: A two-sided rank sum test (comparing differences between samples) yields p=`r rs1` for the mean stock in last period of the game.

@@

Although there clearly is growth, groups do not realize the maximal potential efficiency: under full cooperation, a group can accumulate at maximum `r (80*1.5^10) %>% trunc()` tokens or EUR `r (80*1.5^10 /20) %>% trunc()`. This is depicted in the left panel of @fig-growth-heterogeneity, where one can see the average wealth over time by sample. The panel illustrates for both samples that growth was continuous and surprisingly linear, given the exponential character of the game's design. Despite somewhat differing contribution behavior between samples, neither the eventual wealth nor the corresponding growth paths differed. Differences in contribution behavior did, thus, not translate to significantly different wealth outcomes.

@@

Given the different samples and the possibility of endogenous growth--which essentially is the main feature of the game--I ask whether and how the inequality grows _within_ groups. @fig-gini-time-series illustrates that inequality did grow: at the end of the game, the original and the replication groups exhibit an average Gini coefficient of `r originalGini` and `r replicationGini`, respectively.^[The two-sided rank sum test (comparing differences between samples) yields p=`r rs1` for the mean Gini coefficient in last round of the game.] Because every participant started with the same initial endowment (in _period 0_, so to speak), every group started equally--with a Gini coefficient equaling zero.

@@

**Result 1.** _The `NOPUNISH 10` treatment of GMTV can be replicated because the replication data resemble the original data with respect to initial and final contributions, wealth and growth as well as inequality._

This is remarkable given the different sample and language, the different software and user interface as well as the online setting during the COVID19 pandemic. The result suggests that, by and large, the sum of these factors did not affect people’s preferences towards cooperation.

>>>>

---
title: Revisiting 'Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  I revisit the dynamic public goods game developed by Gächter et al. (2017) to study cooperation under dynamic interdependencies. Collecting data from both a convenient (students) and an inconvenient (general population) sample, I not only reproduce some of the authors' original observations but also test their novel game's generalizability. Appending a charitable dictator game, I find no correlations between behavior in the charitable context and the dynamic game. This applies to students and the general population sample alike. Because the study of inexperienced general population samples raises methodological challenges, such as fatigue and dropouts, this research approaches them. In doing so, I provide simple solutions to run reliable interactive experiments online. Furthermore, this article showcases the use of literate programming and version control which I argue are convenient tools to make pre-registrations more credible and flexible.

@@

Today's actions are tomorrow's result. There are many settings in which current decisions affect future outcomes and with it, future decision spaces. 
<!-- Think of lifetime consumption‐smoothing and save more tomorrow -->
Opting for environmentally friendly policies today not only reduces carbon dioxide emissions immediately but also helps us to reach the Paris climate targets tomorrow. Deferring these policies, may not necessarily prevent us from reaching these targets, but it requires more effort in the future compared to a path that includes immediate action [@HaenselEtAl2022]. Aiming at certain goals, today's actions (or the omission thereof) not only affect intermediate outcomes but also the number of paths one can choose from that lead to that specific goal.

@@

A game designed by Gächter, Mengel, Tsakas & Vostroknutov [-@GMTV2017 hereafter, GMTV] as well as Stefan Große (unpublished), shows that it is fairly simple to add this element of realism. They incorporate interdependencies into a _dynamic_ public goods game (dPGG) by defining endowments as the income of previous periods. Consequently, participants' actions in a given period affect their number of actions in subsequent periods: the more (less) they earn now, the more (less) they can contribute in the next period. Importantly, this modification qualitatively yields the same rational predictions as the static game (i.e. free-riding and the under-provision of the public good). It is thus, equally well suited to study dilemma situations.

Because there is surprisingly little experimental research on interdependencies^[See e.g. @BattagliniEtAl2016, @Rockenbach2017 or @Moser2019.], I reproduced one of GMTVs' treatments to compare dynamics across (in)experienced samples to investigate its generalizability.
@GKLS2020 find that static public goods games do not generalize well to real-world climate action. They also find that generalizability depends on the structural resemblance of the public goods game with the context of climate change mitigation: Greater resemblance improves generalizability. Because GMTVs' dynamic setting has a more realistic property—namely, interdependencies—one would expect it to be better suited to inform public policy. To test this intuition, I not only ran the experiment with different samples but also
<!--, that is, with participants who make their decisions in an environment which is more natural than the lab.-->
observed the participants' behavior in voluntary climate actions (VCA).^[The VCA is a (charitable) dictator game where each participant is a dictator dividing her budget between herself and some organization linked to the reduction of CO2 emissions. @EckelGrossman1996 were the first to implement a charitable dictator game observing contributions of 30% of the endowments. Like @GKLS2020, @CarpenterEtAl2008 report that students make lower contributions to charity than community members.] This yielded a setting similar to the one of @GKLS2020, which allows me to analyze how behavior in the abstract game translates into real-world action across samples. This research shows that the dynamic setting _does not_ add any advances to the generalizability of results.

@AGM2018 conducted static public goods games in the lab and on MTurk to draw (and report) lessons from online experimentation. This study extends this literature [see, e.g., @SnowbergYariv_2021; @GuptaRigottiWilson_2021; @GoodmanPaolacci2017; @AmirEtAl2012] by focusing on an _inconvenient_ sample (i.e. one that is a completely inexperienced sample that has not been exposed to interactive experiments before) playing a computationally more complex game. This required me to design robust (and thus, more complex) software to minimize attrition. I collected para data---which capture screen time and survey navigation, for instance---to assess the desired fluency and feasibility of the experiment. Like @AGM2018, this research is of practical relevance as it reports on the robust design that made the dynamic game feasible.

@@

After commenting on transparent research practices and reporting the methods in @sec-transparency and @sec-methods, this paper is organized along these findings: The confirmatory reproduction is reported in @sec-replication, the online feasibility in @sec-feasibility, and the generalizability in @sec-generalizability. @sec-conclusion concludes. Because I observed some (qualitatively inconsequential) differences in reproducing GMTVs' analysis using their data, I report these in the [Appendix](@sec-appendix).

@@

Well before credibility crises comprised a variety of disciplines^[See e.g., @CamererEtAl2016; @BrodeurEtAl2016; @BrodeurEtAl2020; @ChristensenMiguel2018; @FerraroShukla2020 who focus on economics.], @bemwriting [p. 2] indirectly conceded discrepancies between fuzzy research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to unravel the fuzzy back-and-forth between exploratory and confirmatory research. Prominently, pre-registrations were established to tie the researchers' hands with respect to p-hacking and ex-post theorizing when doing confirmatory research---reducing observable discrepancies between Daryl Bem's above-mentioned (a) and (b).

While pre-registrations could help to make empirical sciences more credible, they come at costs that are discussed by @Olken2015 as well as @Krishna2021, @SimmonsEtAl2021, and @PhamEtAl2021. One important concern is that strict pre-registrations may severely discount any additional analysis of the data inspired by surprise results. In practice, there are at least two ways to approach that problem: First, one could (and should) add unforeseen analyses to the article but label them as exploratory. Second, one could write a minimal pre-registration leaving enough degrees of freedom for the researcher to incorporate such surprises without further notice. The key difference between the two approaches is the deliberate omission of transparency.

For the field of neuropsychopharmacology @WaldronAllen2022 document that the depth of pre-registrations differs indeed: it is spanning from notes on hypotheses only to fully specified, detailed experimental protocols with accompanying power analyses and sampling plans. Treating all pre-registrations equally may thus lead to an adverse effect: If one can acquire the _'pre-registered'_ label with minimal effort and maximum flexibility, the label itself may run into danger to become worthless and to miss the target of making science more credible.

This erosion of credibility can be avoided by the means of transparency---not because transparency can tighten the ties but also because it facilitates scrutiny, policing [see @AnkelPetersFialaNeubauer2023] and forensics: transparent research practices make it easier for others to track and understand a researcher's reaction to surprises in the data, which, in turn, gives the researcher more flexibility to react to these surprises. As additional transparency is associated with additional effort, I propose version control as a natural and easy-to-implement solution augmenting pre-registrations.

Contemporary experimental research is, to a certain degree, a computational science: experiments often require complex software and the analysis thereof can result in complicated scripts. In addition, much research is carried out in collaborative efforts. For these reasons alone, it makes sense to implement version control systems such as GitHub and GitLab as well as OSF (albeit less powerful) that ensure that files are stored consistently and that changes are tracked. When these systems are implemented in the whole research process, one can archive different states of the software and visualize differences between different states. For instance, one can archive the analysis before and after the data was collected or the experimental software used in a pilot and the eventual experiment. Importantly, one can keep projects private and change their visibility to public only after the corresponding article is accepted for publication. Taken together, these features _automatically_ enhance transparency and facilitate scrutiny, policing, and forensics whilst also making software development and collaboration more efficient. As a consequence, the implementation and eventual disclosure of version control come not only with meta-scientific but also with organizational benefits.

The experiment reported in this article qualifies for the _'pre-registered'_ label at first glance:
The analysis was pre-registered in the American Economic Association's RCT Registry [@preregistration]. Further, I pre-registered the exact analyses I planned to run when I designed the experiment on [GitHub](https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R).^[bit.ly/3Pnak5H. To run the code, you need to execute the .Rmd files in this repository in the order that is indicated by file names. Detailed instructions can be found in the README file accompanying the analysis scripts.]

At second glance, the analysis that made the most sense after having collected the data and that is reported here is a little different though: Because the more representative subject pool was exhausted earlier than expected, I recruited students which opened up a new research direction: assessing generalizability. Hence, this article combines both confirmatory as well as exploratory research.

Accordingly, I edited many of the analysis scripts.^[The source code of this document contains the analysis and can be found [here](https://github.com/Howquez/GMTV/blob/main/analysis/article/paper.qmd): https://bit.ly/43LCyLw] However, because the originally planned analysis code is archived and reproducible, these changes are transparent. Importantly, I followed a literate programming approach [@Knuth_1984; @AkhtarYe_2023]. Hence, all documents needed to analyze the data and to write the report stem from the same source. This establishes consistence between the commands one tells the statistical software to do and the explanation one tells human beings one told the statistical software to do. As such, (a), (b), and the transition from (a) to (b) are not only transparent but also comprehensible.

<!--
I archived detailed analyses scripts and created a [branch](https://GitHub.com/Howquez/coopUncertainty/tree/July21Replication) a few days _before_ we collected data and pre-registered the analysis via the American Economic Association's RCT Registry [@preregistration].
-->

@@

In the terminology of @Hamermesh2007, I ran both a _pure_ as well as a _scientific_ reproduction^[In fact, @Hamermesh2007 uses the term _replication_ instead of _reproduction_. The reason I deviate is that I do not compare different treatments with each other to _replicate_ a treatment effect. Instead, I _reproduce_ the results of one specific treatment arm.] of one treatment of GMTVs' dynamic public goods game. The pure reproduction re-analyzes the original data. [Appendix A](#A:-Pure-Replication) documents [the errors I identified]() in the original paper. The scientific reproduction, where I utilize a different sample drawn from a different population in a different situation, is described in the following sections.

@@

Considering the design, the 'only' differences between such a _static_ game (with partner matching) and GMTVs' design are _dynamics_: Instead of receiving fresh endowments every period, participants receive one endowment only at the beginning of the first period. A participant's endowment in the second period is the wealth she accumulated in the first period. A participant's endowment in the third period is the wealth she accumulated in the first two periods. And so on. Hence, a decision in one period has consequences on future endowments and, ultimately, growth paths. For this reason, the game is described as a _dynamic_ public goods game.

@@

Like GMTV I employ a real giving task after the abstract game. Specifically, I employed a (charitable) dictator game where each participant is a dictator dividing her budget between herself and some organization. Whereas GMTV chose _Doctors without Borders_ as a receiver, I chose another organization in another context: like @GKLS2020, I employed a VCA, where participants could donate any amount of their earnings to offset carbon dioxide (that is, retire emission permits from the EU ETS).^[Importantly, @GKLS2020 made the VCA decision with a fresh endowment _before_ they played the abstract game. I deviate from their procedure to match GMTVs' procedure.]
To ensure that each participant had the same basic level of information about the impact of their decision, I provided some basic information about the mechanism. The information also highlighted that the mitigation came into operation on a European level.
Finally, I informed the participants that the documentation of individual and aggregate contributions was to be posted immediately after the conclusion of the sessions online. To avoid privacy or social image concerns, participants learned their unique and random IDs, which they needed to identify their individual contributions.
The document certified that their contributions have been used to offset [1.82 tons](https://www.compensators.org/compensatelist/?searchterm=stefan+traub) of carbon dioxide emissions.

@@

I recruited the participants from the so-called _[HamburgPanel](https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html)_ using HROOT [@hroot]. The panel is provided by the University of Hamburg's Research Laboratory, which used a randomized last digits approach to build the panel while drawing from the population of citizens of Hamburg, Germany. Because the sample was exhausted at one point, I also recruited students from the University of Hamburg.

@@

Throughout this paper, I will compare the results of my experiment with the results of GMTVs' NOPUNISH 10 Period treatments. I am thus, referring to three different samples utilized at two points in time: the University of Nottingham's students (in late 2012), Hamburg's citizens, and the University of Hamburg's students (both in July 2021). @tbl-sample-properties gives an overview.

@@

The experiment was logistically complex for several reasons. First, the sample was inexperienced. Second, the experiment was interactive and synchronous. Third, the underlying game was dynamic and interdependent. This makes dropouts not only more likely but also more expensive, which is why attrition was a major concern in implementing the experiment.

I chose oTree [@oTree] to implement the experiment because it is open-source, well-documented, and very flexible. Its [Bootstrap](https://getbootstrap.com/) (a powerful frontend toolkit) integration allowed me to make the graphical user interface interactive, appealing, and easy to navigate. The [Highcharts library](https://www.highcharts.com/) made it easy to visualize results and communicate dynamics. Insofar, oTree served as a good tool to enhance the participants' user experience and thus, to make dropouts less likely.

Which features were required to handle dropouts? First, participants had to be matched to form a group _after_ comprehension questions were answered successfully. Importantly, participants were grouped by the order they answered these questions to reduce waiting times. While waiting for other players to form the group, the participants saw a wait-page informing them that they are waiting for other participants to arrive and that they do not have to wait for longer than 10 minutes. The screen also informed them that they would receive a _patience bonus_ of one Euro after the expiration of that time (or what was left of it). Second, participants only had 10 minutes to make the first contribution and 4 minutes for the remaining contributions. After this time expired, participants were replaced by bots that made random contributions. In this case, the remaining group members were informed about the replacement. Both features were implemented to limit wait times and boredom for other participants. @sec-feasibility shows that the first feature became effective in some cases, whereas the second feature did not.

@@

Participants entered the experiment at appointed times remotely from home. They first saw a welcome screen. After agreeing to the privacy policy, they could proceed to the instructions individually. Having read these instructions, each participant has also seen a demo-screen explaining the user interface. Before proceeding, they had to answer six comprehension questions correctly to avoid confusion in later stages [@FerraroVossler2010]. Subsequently, they saw a waiting screen until they could be matched with three other participants, who have answered the comprehension questions correctly. Once matched, they were exposed to the decision screen over ten periods. At the end of the last period, participants saw the results of all periods. Subsequently, they made their VCA decision, before I elicited risk preferences [@HoltLaury2002] and finished with GMTVs' questionnaire.

@@

The experiment lasted around `r total_completion_time` minutes on average. The earnings  averaged `r earningsMean %>% round(digits = 2)` Euros (sd = `r earningsSD  %>% round(digits = 2)`).^[This value include earnings from the incentivized risk elicitation task that is not part of the analysis.]

@@

## Reproducibility {#sec-replication}

@@

First, I ask whether the samples differ with respect to their initial contributions to the public good. Is the reproduction sample (consisting of both students and non-students) more pro-social than the original sample? A two-sided rank sum test reveals that it is not (p=`r rs1`). Both samples contributed 10 tokens, that is, 50% of their endowments on average (median and mean). Moreover, both samples' initial contributions resemble the initial contributions participants usually make in the static game with partner matching.^[See Figure 3B in @fehrgaechter2000 [p.989], for instance.] However, in the dynamic game presented here, I am particularly interested in the subsequent periods because differences add up exponentially. Do the two groups remain similar over the course of time?

@@

In particular, do the two samples' contributions follow the same path over the 10 periods they played? The answer is _yes and no_. @fig-share-of-contributions and @tbl-contribution-periods illustrate that the samples make similar contributions at the beginning and the end of the game but behave differently in between. More precisely, the left panel--depicting the average contributions in absolute terms--shows that the original sample contributed more than the reproduction sample _in all but the first and last period_.^[In period five, this difference is significant: a two-sided rank sum test yields p=`r rs2`. ] For this reason, the original sample's behavior differs from the reproduction sample's behavior in two aspects: it contributes more and exhibits a considerable drop in the last period (whereas the reproduction sample's contributions flatten).

@@

Again, both samples' behavior resembles the contributions participants usually make in the static game with partner matching: contributions equal approximately half of the endowments in the very first period and decrease to around ten percent of endowments by the last period.^[The right panel is thus, comparable to the visualizations _and results_ in the static game. See, for instance, Figure 1B in @fehrgaechter2000 [p.986].] In the dynamic game presented here, however, different paths lead to different levels of wealth -- even if they share the same start- and endpoints. I am thus, more interested in the contributions' implications for wealth generation and growth.

@@

How do the different contribution paths translate into wealth?^[To measure wealth and growth, I define a variable called _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied, and redistributed).] Given that the original sample contributed more, one would expect the respective groups to be  more wealthy. A mere mean comparison indicates just that: An average group in the original sample accumulated about `r round(originalWealth)` tokens. In contrast, the average group in the reproduction sample accumulated about `r round(replicationWealth)` tokens. This difference is insignificant at conventional levels though: A two-sided rank sum test (comparing differences between samples) yields p=`r rs1` for the mean stock in the last period of the game.

@@

Although there clearly is growth, groups do not realize the maximal potential efficiency: under full cooperation, a group can accumulate a maximum of `r (80*1.5^10) %>% trunc()` tokens or EUR `r (80*1.5^10 /20) %>% trunc()`. This is depicted in the left panel of @fig-growth-heterogeneity, where one can see the average wealth over time by sample. The panel illustrates for both samples that growth was continuous and surprisingly linear, given the exponential character of the game's design. Despite somewhat differing contribution behavior between samples, neither the eventual wealth nor the corresponding growth paths differed. Differences in contribution behavior did, thus, not translate to significantly different wealth outcomes.

@@

Given the different samples and the possibility of endogenous growth---which essentially is the main feature of the game---I ask whether and how the inequality grows _within_ groups. @fig-gini-time-series illustrates that inequality did grow: at the end of the game, the original and the replication groups exhibit an average Gini coefficient of `r originalGini` and `r replicationGini`, respectively.^[The two-sided rank sum test (comparing differences between samples) yields p=`r rs1` for the mean Gini coefficient in the last round of the game.] Because every participant started with the same initial endowment (in _period 0_, so to speak), every group started equally--with a Gini coefficient equaling zero.

@@

**Result 1.** _The `NOPUNISH 10` treatment of GMTV can be replicated because the replication data resemble the original data with respect to initial and final contributions, wealth, and growth as well as inequality._

This is remarkable given the different sample and language, the different software and user interface as well as the online setting during the COVID-19 pandemic. The result suggests that by and large, the sum of these factors did not affect people’s preferences towards cooperation.
